Exploring probability distributions for bivariate temporal granularities

### 2019-11-17 main.Rmd
harmony and clash section revised

Now, suppose we want a cyclic representation of two periodic granularities.

@@

# Synergy of time granularities

@@

Before exploring the behavior of a "dependent variable" across time granularities,
it is important to know how these granularities interact with each other. If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable. To define a general framework, we define harmony and clashes. Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in a non-compatible ways and hinder the process of exploration of the dependent variable if plotted together.

@@

Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps row numbers to a set $\{B_1,B_2,B_3,\dots, B_m$\}.  That is, let $S_{ij}$ be the set of row numbers such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$. Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$. The graphs that don't work are those where many of these 12 sets are empty. Now, many situations can lead to any of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty.

Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
These are called "structurally" empty combinations. Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month. Here $C_1$ can take 31 values while $C_2$ can take 5 values. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month. In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. These are structurally empty sets in that it is impossible for them to have any observations.

@@

An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year. Here $C_1$ can take 7 values while $C_2$ can take 12 values. So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY. All of these are non-empty because every DOW can occur in every month. So graphics involving $C_1$ and $C_2$ are potentially useful.

Therefore, combinations of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
And the ones that do not lead to any missing combinations are called **harmonies** as they promote the exploratory analysis through visualization.

>>>>

Now, suppose we want a cyclic representation of two periodic granularities G and H, where H is periodic with respect to G.

@@

# Synergy of cyclic time granularities

@@

Before exploring the behavior of a "dependent variable" across time granularities,
it is important to know how these granularities interact with each other. If two time granulaties of interest interact, the relationship between each of the interacting variables and the dependent variable will depend on the value of the other interacting variable. To define a general framework, we define harmony and clashes.

<!-- Harmonies are the pair of granularities that aid the process of exploratory analysis and clashes are pair that interact in non-compatible ways and hinder the process of exploration of the dependent variable. -->

@@

Suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A_1,A_2,A_3,\dots, A_n$\}, and $C_2$ maps index set to a set $\{B_1,B_2,B_3,\dots, B_m$\}.  That is, let $S_{ij}$ be the set of index set such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$. Since, $C_1$ has n levels and $C_2$ has m levels there will be $nm$ such sets $S_{ij}$.Now, many situations can lead to any of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty.

Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
These are called "structurally" empty combinations. Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month. Here $C_1$ can take 31 values while $C_2$ can take 5 values. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month. These are structurally empty sets in that it is impossible for them to have any observations.

<!--  In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. -->

@@

An example when there will be no empty combinations could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year. Here $C_1$ can take 7 values while $C_2$ can take 12 values. So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY. All of these are non-empty because every DOW can occur in every month.

Therefore, pair of circular/aperiodic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**.
And the ones that do not lead to any missing combinations are called **harmonies**.

<!-- as they promote the exploratory analysis through visualization. -->
